Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Verification

### 2015-02-23 documentation/userguide.tex
Added documentation about problem setup and parameters.

\clearpage

@@

>>>>

@@

\chapter{Merger Problem}

This chapter describes the setup of the white dwarf merger problem. It includes descriptions of 
the physics involved, the code options and runtime parameters, the diagnostic output 
and other I/O information, and the analysis routines. A full description of the physics 
and the associated numerical algorithms is included in the CASTRO code paper \cite{castro}.

\section{Physical Description}

We model two white dwarfs in a rotating binary system. The white dwarfs are composed of 
a mixture of elements that is principally carbon and oxygen, and are individually supported 
against gravity by degenerate electron material at non-zero temperature. This matter is described well 
for arbitrary degeneracy and relativity levels by the Helmholtz equation of state \cite{timmes_swesty:2000}.
It includes ideal gas contributions from the ions, a simple blackbody recipe for the contribution from 
the radiation field, and optionally Coulomb corrections.

CASTRO includes full self-gravity with the Poisson equation, as well as the ability to perform a
simulation in a rotating reference frame with fixed rotation period. By default the $z$ axis will 
be the rotation axis, with the stars initially aligned along the $x$ axis. We select the 
orbital period of the system, and the masses of the \textit{primary} (more massive) and 
\textit{secondary} (less massive) white dwarfs. We then use Kepler's third law to determine the 
appropriate initial orbital distance. The center of mass of the system corresponds with the center 
of the domain, which is the origin; the domain is $1.024 \times 10^{10}\ \text{cm}$ wide in each 
coordinate direction, so that each dimension runs from $-5.12 \times 10^{9}\ \text{cm}$ to 
$+5.12 \times 10^{9}\ \text{cm}$. In a rotating reference frame, we set the stars at rest. 
We can instead run a simulation in the inertial reference frame. To make it equivalent, we
start all zones in the domain with a velocity corresponding to the equivalent rotating 
reference frame velocity. This means that the stars will be synchronously rotating (though we
may relax this assumption in future work).

At present nuclear reactions are not enabled, but they will be soon.

\section{Runtime Parameters}

There are a few CASTRO input parameters that you will typically want to control when running problems. 
Among the most major is whether we are doing rotation (\texttt{castro.do\_rotation}) and if so, 
what the rotational period is (\texttt{castro.rotational\_period}). You'll also want to change the 
\texttt{stop\_time}, which is when the simulation will terminate. You should also avoid flooding 
yourself with plotfiles and checkpoints by setting \texttt{amr.plot\_files\_output} and 
\texttt{amr.checkpoint\_files\_output}, respectively; these are the simulation time differences
between successive plotfiles and checkpoints. Note that CASTRO does not adjust the timestep 
to meet this criterion, so it will generate these at the first timestep that passes this interval.

The \texttt{wdmerger} parameters of interest are all controlled through the \texttt{probin} file. 
These control the layout of the stars on the grid, the tagging options for refinement, and some other things. 
A \texttt{probin} file is composed of three Fortran namelists; a namelist is enclosed in 
\texttt{\&namelist ... /}. The \texttt{fortin} namelist controls the main system parameters. 
In the following list we describe the parameters; you can see their current defaults by 
examining \texttt{source/probin}.
\begin{itemize}
  \item \texttt{mass\_P}: mass of the primary white dwarf, in solar masses (real)
  \item \texttt{mass\_S}: mass of the secondary white dwarf, in solar masses (real; if less than zero, 
    then no secondary star will be generated, and the primary star will be loaded onto the grid at the 
    center of the domain)
  \item \texttt{nsub}: number of sub-intervals to divide a zone into when interpolating from the initial model (integer)
  \item \texttt{orbital\_kick}: whether or not to kick the stars with an orbital velocity corresponding to their 
    Keplerian rotation rate (logical; this only makes sense in the inertial reference frame with \texttt{castro.do\_rotation = 0})
  \item \texttt{damping}: whether or not to exert a constant damping force on the orbit (logical)
  \item \texttt{damping\_alpha}: strength of damping force (real)
  \item \texttt{bulk\_velx}: uniform velocity to give to all gas in the domain in the $x$-direction (real)
  \item \texttt{bulk\_vely}: uniform velocity to give to all gas in the domain in the $y$-direction (real)
  \item \texttt{bulk\_velz}: uniform velocity to give to all gas in the domain in the $z$-direction (real)
  \item \texttt{interp\_temp}: whether or not to take temperature as given when interpolating from our initial model (logical;
    if false, we will interpolate based on pressure)
  \item \texttt{maxTaggingRadius}: fractional size of the domain outside of which we won't perform any refinement (real)
\end{itemize}

The \texttt{tagging} namelist is described in the AMR chapter of the CASTRO user's guide. At present, we 
only tag based on absolute values of the density and temperature.

The \texttt{extern} namelist contains parameters that are used by the microphysics modules. In particular, 
the equation of state usually relies on parameters passed in this way. For the Helmholtz EOS, the relevant 
parameters are \texttt{eos\_input\_is\_constant}, which determines whether or not we update the internal 
energy on an EOS call, and \texttt{use\_eos\_coulomb}, which determines whether we include Coulomb 
interactions in the EOS evaluation.
